Normalization

### 2020-05-04 main_figures/clustering.Rmd
updated many figures to improve readability...

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(pipeComp)
})
source("../misc_functions.R")
theme_set(theme_cowplot(font_size = 12))
```

@@

tmp <- scale(pipeComp:::.prepRes(cl, "ARI", agg.by=c("norm","graph","clustmethod"))) + 
  scale(pipeComp:::.prepRes(cl, "MI", agg.by=c("norm","graph","clustmethod")))
o <- names(sort(rowMeans(tmp, na.rm=TRUE)+apply(tmp,1,na.rm=TRUE,FUN=median), decreasing = TRUE))
o <- gsub("norm=norm\\.","",o)
o <- c(o[grep("seuratvst",o)],o[grep("seuratvst",o,invert=TRUE)])
h <- scrna_evalPlot_clust(cl, what="MI", agg.by=c("norm","graph","clustmethod"), agg.fn = median, scale=TRUE, anno_legend = FALSE, reorder_rows=o, row_names_max_width = unit(8, "cm"), row_names_gp = gpar(fontsize = 11)) +
  scrna_evalPlot_clust(res, what="ARI", agg.by=c("norm","graph","clustmethod"), agg.fn = median, scale=TRUE, reorder_rows = o, anno_legend = FALSE) +
  scrna_evalPlot_clust(res, what="min_pr", agg.by=c("norm","graph","clustmethod"), agg.fn = median, scale=TRUE, reorder_rows = o, anno_legend = FALSE) +
  scrna_evalPlot_clust(res, what="ARI", atTrueK = TRUE, agg.by=c("norm","graph","clustmethod"), agg.fn = median, reorder_rows = o, anno_legend = FALSE, row_names_max_width = unit(8, "cm"), row_names_gp = gpar(fontsize = 11))
h <- pipeComp:::.renameHrows(h, renameFn)
h2 <- grid.grabExpr(draw(h))
```

```{r clustering_time}
clo$elapsed <- res$elapsed$stepwise$clustering$elapsed
# keep reasonable parameters for a realistic time
res$evaluation$clustering <- clo[which(clo$steps>2 & clo$k <=20),]
h3 <- scrna_evalPlot_clust(res, what="elapsed", agg.by=c("graph","clustmethod"),
                           anno_legend = TRUE, scale=TRUE, value_format = "%.1f" )
h3 <- h3 + scrna_evalPlot_clust(res, what="nbClust.diff", 
                                agg.by=c("graph","clustmethod"), agg.fn = median,
                                reorder_rows=h3, scale = FALSE, value_format = "%.1f", 
                                col = circlize::colorRamp2(c(-5, 0, 5), c("red", "white", "Darkblue")),
                                value_cols = c("black","black"), 
                                row_names_max_width = unit(8, "cm"),
                                row_names_gp = gpar(fontsize = 12) )
h3 <- pipeComp:::.renameHrows(h3, renameFn)
h3 <- grid.grabExpr(draw(h3))
```

```{r clustering, fig.width=12, fig.height=8}
pdf("clustering.pdf", width=15, height=10)
plot_grid(h2, h3, labels=LETTERS[1:2], ncol=1, rel_heights = c(7,4))
dev.off()
plot_grid(h2, h3, labels=LETTERS[1:2], ncol=1, rel_heights = c(7,4))
```
Clustering accuracy (A), computing time (B left) and mean deviation from the true number of clusters (B right). The (exact) rank-based SNN graph represents the default setting in `scran`-based clustering; however we find that the Annoy approximation to the KNN diminishes the computing costs without negatively affecting the accuracy.

>>>>

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(pipeComp) # version >= 0.99.26
})
source("../misc_functions.R")
theme_set(theme_cowplot(font_size = 12))
```

@@

tmp <- scale(pipeComp:::.prepRes(cl, "ARI", agg.by=c("norm","graph","clustmethod"))) + 
  scale(pipeComp:::.prepRes(cl, "MI", agg.by=c("norm","graph","clustmethod")))
o <- names(sort(rowMeans(tmp, na.rm=TRUE)+apply(tmp,1,na.rm=TRUE,FUN=median), decreasing = TRUE))
o <- gsub("norm=norm\\.","",o)
o <- c(o[grep("seuratvst",o)],o[grep("seuratvst",o,invert=TRUE)])

h <- renameHrows(
  evalHeatmap(res, step="clustering", what=c("MI","ARI","min_pr"), row_split = norm, font_factor = 0.85,
              agg.by=c("norm","graph","clustmethod"), anno_legend=FALSE) +
  evalHeatmap(res, step="clustering", what="ARI", agg.by=c("norm","graph","clustmethod"), font_factor = 0.85, filter=n_clus==true.nbClusts, title="ARI at\ntrue k", name="ARIatk",anno_legend=FALSE), function(x) sapply(strsplit(x," > "), FUN=function(x) paste(x[-1],collapse=" > ")))

h2 <- grid.grabExpr(draw(h))
```

```{r clustering_time}
clo$elapsed <- res$elapsed$stepwise$clustering$elapsed
# keep reasonable parameters for a realistic time
res$evaluation$clustering <- clo[which(clo$k<=20),]
h3 <- evalHeatmap(res, step="clustering", what="elapsed", agg.by=c("graph","clustmethod"), font_factor = 0.85,
                           anno_legend = TRUE, scale=TRUE, value_format = "%.1f", reorder_rows=TRUE,
                  row_names_max_width = unit(8, "cm"))
h3 <- h3 + evalHeatmap(res, step="clustering", what="nbClust.diff", 
                       row_names_max_width = unit(8, "cm"),
                        agg.by=c("graph","clustmethod"), agg.fn = median, font_factor = 0.85,
                        reorder_rows=h3, scale = FALSE, value_format = "%.1f", 
                        col = circlize::colorRamp2(c(-5, 0, 5), c("red", "white", "Darkblue")),
                        value_cols = c("black","black") )
h3 <- grid.grabExpr(draw(h3))
```

```{r clustering, fig.width=11, fig.height=8}
pdf("clustering.pdf", width=11.5, height=8)
plot_grid(h2, h3, labels=LETTERS[1:2], ncol=1, rel_heights = c(7,4))
dev.off()
plot_grid(h2, h3, labels=LETTERS[1:2], ncol=1, rel_heights = c(7,4))
```
Clustering accuracy (A), computing time (B left) and mean deviation from the true number of clusters (B right). The (exact) rank-based SNN graph represents the default setting in `scran`-based clustering; however we find that the Annoy approximation to the KNN diminishes the computing costs without negatively affecting the accuracy.

### 2020-05-04 main_figures/dimensionality.Rmd
updated many figures to improve readability...

```{r, include=FALSE}
#devtools::load_all("../../pipeComp_biorxiv/pipeComp/")
```

@@

```{r, fig.width=8, fig.height=4}
res1 <- readRDS("../data/results_dimensionality.rds")
res2 <- readRDS("../data/results_dimensionality2.rds")
sw1 <- res1$dimensionality$clust.avg.silwidth$selected
sw2 <- res2$dimensionality$clust.avg.silwidth$selected
sw2 <- sw2[grep("Local\\.b",row.names(sw2),invert = TRUE),] # redundant
sw <- rbind(sw1,sw2[setdiff(row.names(sw2),row.names(sw1)),])
sw <- sw[grep("=doublet",row.names(sw)),]
n <- sapply(strsplit(row.names(sw),"="),FUN=function(x) rev(x)[1])
w <- grep("seuratvst",row.names(sw))
n[w] <- paste0("sctransform+",n[w])
row.names(sw) <- n
sw2 <- sw[order(grepl("sctransform",row.names(sw))*1000+rowMeans(sw)+apply(sw,1,probs=0.2,FUN=quantile),decreasing=TRUE),order(colMeans(sw),decreasing=TRUE)]
h1 <- Heatmap(sw2, cluster_columns = FALSE, cluster_rows = FALSE, 
        name="silhouette width", col=pipeComp:::.silScale(sw2), show_column_names = FALSE,
        column_title = "Subpopulation mean silhouette width",
        bottom_annotation=pipeComp:::.ds_anno(colnames(sw2)), heatmap_legend_param = list(direction="horizontal", legend_width=unit(3, "cm")))
p2 <- grid.grabExpr(draw(h1, heatmap_legend_side="bottom"))

cl1 <- res1$clustering
cl2 <- res2$clustering
res2 <- rbind(cl1, cl2[setdiff(row.names(cl2),row.names(cl1)), colnames(cl1)])
res2 <- res2[grep("=doublet",row.names(res2)),]
method <- as.character(parsePipNames(row.names(res2))$dims)
w <- grep("seuratvst", row.names(res2))
method[w] <- paste0("sctransform+",method[w])
row.names(res2) <- paste0("method=",method,";",row.names(res2))

h2 <- scrna_evalPlot_clust(res2, what="NMI", agg.by = c("method"), scale=TRUE, reorder_rows = TRUE, anno_legend = FALSE)
h2 <- h2 + 
  scrna_evalPlot_clust(res2, what="min_pr", agg.by = c("method"), reorder_rows=h2, scale=TRUE, anno_legend = FALSE) + 
scrna_evalPlot_clustAtTrueK(res2, agg.by = c("method"), reorder_rows=h2, scale=TRUE, anno_legend = FALSE)
p3 <- grid.grabExpr(draw(h2))
```

>>>>

```{r, include=FALSE}
devtools::load_all("../../pipeComp_biorxiv/pipeComp/")
source("../../pipeComp_biorxiv/pipeComp/inst/extdata/scrna_alternatives.R")
```

@@

```{r, include=FALSE}
# the second part of these results was generated with pipeComp >0.99.26
devtools::load_all("../../pipeComp/")
```

```{r, fig.width=8, fig.height=4}
res <- readRDS("../data/results_dimensionality.rds")
H <- renameHrows(
  scrna_evalPlot_silh( res, step="dimensionality", reorder_rows=TRUE, 
                       row_split=gsub("norm\\.","",norm),
                       heatmap_legend_param=list(direction="horizontal")),
  function(x) gsub("norm\\.","",x))
p2 <- grid.grabExpr(draw(H, heatmap_legend_side="bottom"))

p3 <- renameHrows(
  evalHeatmap( res, what=c("NMI","min_pr"), agg.by=c("norm","dims"), 
                   row_split=gsub("norm\\.","",norm), reorder_rows=TRUE, anno_legend=FALSE) + 
  evalHeatmap(res, what=c("ARI"), agg.by=c("norm","dims"), row_split=gsub("norm\\.","",norm), 
              reorder_rows=TRUE, filterExpr=n_clus==true.nbClusts, 
              anno_legend=FALSE, title="ARI at\ntrue k"),
  function(x) gsub("norm\\.","",x)
)
p3 <- grid.grabExpr(draw(p3))
```

### 2020-05-04 main_figures/doublets.Rmd
updated many figures to improve readability...

pw <- function(i, dot.size=4, ...){
  scores <- res[[i]][,intersect(names(scoresNames),colnames(res[[i]]))]
  colnames(scores) <- scoresNames[colnames(scores)]
  plotROCs(scores, factor(res[[i]]$demuxlet_cls, c("SNG","DBL")), dot.size=dot.size) +
    ylim(0,1) + scale_color_manual(values=cols)
}

p1 <- pw(1) + theme(legend.position="none")
p2 <- pw(2) + theme(legend.position="none")
p3 <- pw(3) + theme(legend.position=c(0.25,0.25))
```

@@

```{r}
mis <- readRDS("../data/misclassification.rds")
lf <- list.files("../datasets/", full.names = TRUE)
lf <- lf[grep("simMix",lf,invert=TRUE)]
names(lf) <- sapply(strsplit(basename(lf),"\\."), FUN=function(x) x[1])
CDs <- lapply(lf, FUN=function(x) colData(readRDS(x)))
dbl <- lapply(CDs, FUN=function(x) x[,c("pldf.ratio","pldf.class")])
dbl <- dbl[intersect(names(dbl),names(mis))]
for(n in names(dbl)){
  dbl[[n]]$pct_misclassified_cor <- mis[[n]]$corrected_pct_mis[row.names(dbl[[n]])]
  dbl[[n]]$pct_misclassified <- mis[[n]]$pct_misclassified[row.names(dbl[[n]])]
}
d <- as.data.frame(data.table::rbindlist(lapply(dbl,as.data.frame), idcol="dataset"))
d$pldf.class <- factor(d$pldf.class, c("doublet","singlet"))
levels(d$pldf.class) <- c("DBL","SNG")
p4 <- ggplot(d, aes(pldf.class, pct_misclassified, fill=pldf.class)) + geom_violin() + facet_wrap(~dataset, nrow = 1) + theme(axis.title.x=element_blank(), legend.position="none") + ylab("% misclassified") + stat_summary(fun.y="mean", geom="point")
```

```{r figure1, fig.height=6, fig.width=7}
p1m <- plot_grid(p1, p2, p3, nrow=1)
pdf("doublets.pdf", width=6, height=7)
plot_grid(p1m, pt, p4, rel_heights=c(3.5,2,2), labels = LETTERS[1:3], ncol = 1, hjust=0, scale=0.95)
dev.off()
plot_grid(p1m, pt, p4, rel_heights=c(3.5,2,2), labels = LETTERS[1:3], ncol = 1, hjust=0)
```

>>>>

pw <- function(i, dot.size=4, ...){
  scores <- res[[i]][,intersect(names(scoresNames), colnames(res[[i]]))]
  colnames(scores) <- scoresNames[colnames(scores)]
  plotROCs(scores, factor(res[[i]]$demuxlet_cls, c("SNG","DBL")), dot.size=dot.size) +
    ylim(0,1) + scale_color_manual(values=cols) + 
    facet_wrap(as.formula(paste0('~"', names(res)[i], '"')))
}

p1 <- pw(1) + theme(legend.position="none")
p2 <- pw(2) + theme(legend.position="none") + ylab("")
p3 <- pw(3) + theme(legend.position=c(0.25,0.25)) + ylab("")
```

@@

```{r}
mis <- readRDS("../data/misclassification.rds")
#lf <- list.files("../datasets/", full.names = TRUE)
#lf <- lf[grep("simMix",lf,invert=TRUE)]
#names(lf) <- sapply(strsplit(basename(lf),"\\."), FUN=function(x) x[1])
#CDs <- lapply(lf, FUN=function(x) colData(readRDS(x)))
dbl <- lapply(CDs[grep("simMix", names(CDs), invert=TRUE)], 
              FUN=function(x) x[,c("pldf.ratio","pldf.class")])
dbl <- dbl[intersect(names(dbl),names(mis))]
for(n in names(dbl)){
  dbl[[n]]$pct_misclassified_cor <- mis[[n]]$corrected_pct_mis[row.names(dbl[[n]])]
  dbl[[n]]$pct_misclassified <- mis[[n]]$pct_misclassified[row.names(dbl[[n]])]
}
d <- as.data.frame(data.table::rbindlist(lapply(dbl,as.data.frame), idcol="dataset"))
d$pldf.class <- factor(d$pldf.class, c("doublet","singlet"))
levels(d$pldf.class) <- c("DBL","SNG")
p4 <- ggplot(d, aes(pldf.class, pct_misclassified, fill=pldf.class)) + geom_violin() + facet_wrap(~dataset, nrow = 1) + theme(axis.title.x=element_blank(), legend.position="none") + ylab("% misclassified") + stat_summary(fun.y="mean", geom="point")
```

```{r figure1, fig.height=6, fig.width=8}
p1m <- plot_grid(p1, p2, p3, nrow=1, rel_widths = c(9,10,10))
pdf("doublets.pdf", width=8, height=6)
plot_grid(p1m, pt, p4, rel_heights=c(3.5,2,2), labels = LETTERS[1:3], ncol = 1, hjust=0, scale=0.95)
dev.off()
plot_grid(p1m, pt, p4, rel_heights=c(3.5,2,2), labels = LETTERS[1:3], ncol = 1, hjust=0)
```

### 2020-05-04 main_figures/dr.Rmd
updated many figures to improve readability...

---
title: "Figure 5 dimensionality reduction"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
---

@@

```{r dimred, fig.width=12, fig.height=6}
res <- readRDS("../data/results_dimred.rds")
x <- res$dimreduction$clust.avg.silwidth
x <- x[grep("norm\\.seurat\\.noscale.*denoise", row.names(x), invert=TRUE),]
x <- x[order(colSums(apply(x,1,prob=c(0.05,0.5),FUN=quantile))+
                         rowMeans(x), decreasing=TRUE),order(colMeans(x), decreasing=TRUE)]
row.names(x) <- gsub("norm\\.","",apply(parsePipNames(row.names(x))[,c("norm","dr")],1,collapse=" + ",FUN=paste))
row.names(x) <- gsub("seuratvst","sctransform",row.names(x))
h1 <- Heatmap(x, cluster_columns = FALSE, cluster_rows = FALSE, 
        name="silhouette width", col=pipeComp:::.silScale(x), show_column_names = FALSE,
        column_title = "Subpopulation mean silhouette width", row_names_max_width = unit(9,"cm"),
        bottom_annotation=pipeComp:::.ds_anno(colnames(x)), heatmap_legend_param = list(direction="horizontal", legend_width=unit(3, "cm")))
p1 <- grid.grabExpr(draw(h1, heatmap_legend_side="bottom"))

res$clustering <- res$clustering[grep("norm\\.seurat\\.noscale.*denoise", row.names(res$clustering), invert=TRUE),]
h <- scrna_evalPlot_clust(res, what="MI", agg.by=c("norm","dr"), scale=TRUE, anno_legend = FALSE)
h <- h + scrna_evalPlot_clust(res, what="mean_F1", agg.by=c("norm","dr"), scale=TRUE, reorder_rows = h, anno_legend = FALSE) +
  scrna_evalPlot_clust(res, what="min_pr", agg.by=c("norm","dr"), scale=TRUE, reorder_rows = h, anno_legend = FALSE) +
  scrna_evalPlot_clustAtTrueK(res, agg.by=c("norm","dr"), reorder_rows = h, anno_legend = FALSE)
h <- pipeComp:::.renameHrows(h, function(x) gsub("; dr=", " + ", gsub("seuratvst", "sctransform", x)))

p2 <- grid.grabExpr(draw(h))

plot_grid(p1,p2, nrow=2, labels=LETTERS[1:2])
```

**A:** Average silhouette width per subpopulation according to various normalization and dimension reductions. **B:** Clustering accuracy , mean proportion of the variance in the first 5 components explained by real subpopulations (center), and median ARI of the resulting clustering (right).

>>>>

---
title: "Figure 5 dimensionality reduction"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

@@

```{r, include=FALSE}
devtools::load_all("../../pipeComp/")
```

```{r dimred, fig.width=10.5, fig.height=6}
res <- readRDS("../data/results_dimred2.rds")
H <- renameHrows(
  scrna_evalPlot_silh( res, reorder_rows=TRUE, 
                       heatmap_legend_param=list(direction="horizontal")),
  function(x) gsub("norm\\.","",x))
p1 <- grid.grabExpr(draw(H, heatmap_legend_side="bottom"))

p2 <- renameHrows(
  evalHeatmap( res, what=c("MI","min_pr"), agg.by=c("norm","dr"), reorder_rows=TRUE, anno_legend=FALSE,
               title=c("MI","Minimum precision")) + 
  evalHeatmap(res, what=c("ARI"), agg.by=c("norm","dr"), filterExpr=n_clus==true.nbClusts, 
              anno_legend=FALSE, title="ARI at true k"),
  function(x) gsub("norm\\.","",x)
)
p2 <- grid.grabExpr(draw(p2))

pdf("dr.pdf", width = 10.5, height=6)
plot_grid(p1,p2, nrow=2, labels=LETTERS[1:2])
dev.off()
plot_grid(p1,p2, nrow=2, labels=LETTERS[1:2])
```

Silhouette width (**A**) per subpopulation and clustering accuracy (**B**) according to various normalization and dimension reductions.

### 2020-05-04 main_figures/normalization.Rmd
updated many figures to improve readability...

(Requires `pipeComp` version >=0.99.3)

@@

```{r}
p1 <- scrna_evalPlot_DR(res, "log10_total_counts", agg.by = "norm", anno_legend=FALSE)
p1 <- p1 + scrna_evalPlot_DR(res, "log10_total_features", agg.by = "norm", anno_legend=FALSE, reorder_rows=p1)
p1 <- grid.grabExpr(draw(renameNorm(p1)))

p2 <- grid.grabExpr(draw(
  renameNorm(scrna_evalPlot_DR(res, what="meanSilWidth", agg.by = "norm", value_format = "", show_heatmap_legend = TRUE))))

H <- scrna_evalPlot_clust(res, "MI", agg.by = "norm", scale = TRUE, reorder_rows=TRUE, anno_legend=FALSE)
H <- H + scrna_evalPlot_clust(res, "ARI", agg.by = "norm", scale = TRUE, reorder_rows = row.names(H@matrix), anno_legend=FALSE) +
  scrna_evalPlot_clust(res, "ARI", atTrueK = TRUE, agg.by = "norm", scale = TRUE, reorder_rows = row.names(H@matrix), anno_legend=FALSE)
p3 <- grid.grabExpr(draw(renameNorm(H)))
```

```{r figure3, fig.width=9, fig.height=11}
pdf("normalization.pdf", width=10, height=11)
plot_grid(p1,p2,p3, nrow = 3, labels=LETTERS[1:3], rel_heights = c(10,9,10))
dev.off()
plot_grid(p1,p2,p3, nrow = 3, labels=LETTERS[1:3], rel_heights = c(10,9,10))
```

Evaluation of normalization procedures. **A:** Variance in the first principal component explained by 
library size and detection rate after accounting for subpopulation differences. The values are difference in adjusted _R^2_ between a `PC1~subpopulation+covariate` model and `PC1~subpopulation`. **B:** Average silhouette width per true subpopulation, where higher silhouette width means a higher separability. **C:** Clustering accuracy, measured by the mutual information (MI), the adjusted Rand Index (ARI), and the ARI at the true number of cluster.

>>>>

(Requires `pipeComp` version >=0.99.26)

@@

```{r}
classifyNorms <- function(x){
  y <- rep("other", length(x))
  y[grep("sctransform",x)] <- "sctransform"
  y[grep("seurat",x)] <- "seurat"
  y[grep("scran",x)] <- "scran"
  y[grep("stableG",x)] <- "stableG"
  y
}

p1 <- evalHeatmap(res, step="dimreduction", what=c("log10_total_counts","log10_total_features"), 
                  what2="PC1.covar.adjR2", agg.by = "norm", reorder_rows = TRUE, 
                  row_split = classifyNorms(norm), row_title="Variance in PC1 explained by covariate")
lgd <- Legend(col_fun=colorRamp2(1:100, viridisLite::inferno(100)), at=c(1,50,100), 
              title="MADs", labels=c("lowest","median","highest"))
p1 <- grid.grabExpr(draw(renameNorm(p1), merge_legend=TRUE, annotation_legend_list=lgd))

p2 <- grid.grabExpr(draw(renameNorm(
  scrna_evalPlot_silh(res, reorder_rows = TRUE, row_split = classifyNorms(norm),
                      row_title="Subpopulations silhouette width", anno_legend=FALSE)
  ), annotation_legend_side="right"))

lgd <- Legend(col_fun=colorRamp2(1:100, viridisLite::inferno(100)), at=c(1,50,100), 
              title="MADs", labels=c("worst","median","best"), direction="horizontal")
p3 <- grid.grabExpr(draw(
  renameNorm(evalHeatmap(res, step="clustering", what=c("MI","ARI"), agg.by = "norm", 
              reorder_rows=TRUE, anno_legend=FALSE, row_split=classifyNorms(norm), row_title="Clustering accuracy") + 
    evalHeatmap(res, step="clustering", what="ARI", agg.by = "norm", 
                filter=n_clus==true.nbClusts, reorder_rows=TRUE, anno_legend=FALSE,
                row_split=classifyNorms(norm), name="ARI at true k", title="ARI at true k")),
  heatmap_legend_side = "bottom", merge_legend=TRUE, annotation_legend_list=lgd))
```

```{r figure3, fig.width=9, fig.height=11}
pdf("normalization.pdf", width=10, height=11)
  plot_grid(p1,p2,p3, nrow = 3, labels=LETTERS[1:3], rel_heights = c(9,9,10))
dev.off()
plot_grid(p1,p2,p3, nrow = 3, labels=LETTERS[1:3], rel_heights = c(9,9,10))
```

Evaluation of normalization procedures. **A:** Variance in the first principal component explained by 
library size and detection rate after accounting for subpopulation differences. The values are difference in adjusted _R^2_ between a `PC1~subpopulation+covariate` model and `PC1~subpopulation`. **B:** Minimum and mean silhouette width per true subpopulation, where higher silhouette width means a higher separability. **C:** Clustering accuracy, measured by the mutual information (MI), the adjusted Rand Index (ARI), and the ARI at the true number of cluster. The coloring scheme in panels **A** and **B** is based on the per-metric number of median absolute deviations (MADs) from the per-column medians, and mapped to color using a square-root transformation to minimize the effect of outliers. The values shown in the cells are the direct metric values (no transformation except rounding).
